---
title: "Introduction to re2"
output: rmarkdown::html_vignette
description: >
  Start here if this is your first time using re2. You'll find a brief
  introduction to Google's RE2 regular expression matching library,
  and you'll learn about important functions to match, extract, and
  replace patterns in strings.
vignette: >
  %\VignetteIndexEntry{Introduction to re2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(re2)
```

## re2: R interface to the Google's RE2 (C++) regular-expression library

re2 provides functions for matching, replacement, and extraction of
regular expression pattern in a character vector. It can be a drop-in
replacement for other regular expression libraries.

Why use re2?

Perl, PCRE, Python, Ruby, Java, and many other languages rely on
regular expression implementations based on recursive backtracking
that can be excruciatingly slow. The time complexity can be
exponential in some matching situations. In contrast, regular
expression matching using finite automata-based techniques (like re2)
can be simple and fast, guaranteeing linear time execution and a fixed
stack footprint. See references for more information.

### Primary re2 functions

re2 supports four types of operations: matching, substring
    extraction, replacement, and extraction with replacement.

#### re2_matchl() or re2_match()

*Matching* is provided by `re2_matchl()` (or `re2_match()` with logical option
set to TRUE). Matching regexp "(foo)|(bar)baz" on "barbazbla" will
return TRUE, for instance.

```{r}
re2_match("foobar", "foo")

```

*Substring extraction* is provided by `re2_match()`.
Matching regexp "(foo)|(bar)baz" on "barbazbla" will return
submatches '.0' = "barbaz", '.1' = NA, and '.2' = "bar". '.0' is
the entire matching text. '.1' is the first group, and so
on. Groups can also be named.

...

*Replacement* functionality is provided by `re2_replace()` and
`re2_global_replace()`. Replacing first occurrence of
pattern "b+" using replacement string "d" on text "yabba dabba doo"
will result in "yada dabba doo". Replacing globally will result in
"yada dada doo".

...

*Extraction with replacement* is provided by
`re2_extract()`. It is like replacement except that
non-matching text is ignored. Extracting "(.*)@([^.]*)" from text
"bunny@wunnies.pl", using replacement string "\\2!\\1" will result
in "wunnies!bunny". "\\1" and "\\2" are names of capturing
subgroups.

...

In all the above functions regexp patterns can be precompiled and
reused for efficiency. See \code{\link{re2_re2}}. In addition, both
input parameter and result are vectorized.

...

There are also a few utility functions (not needed for most applications):
`re2_get_options()`,
`re2_capturing_group_names()`,
`re2_named_capturing_groups()`,
`re2_number_of_capturing_groups()`,
`re2_quote_meta()`,
`re2_max_submatch()`,
`re2_possible_match_range()`,
`re2_check_rewrite_string()`.

### Syntax


### References

* [Regular Expression Matching Can Be Simple And Fast](https://swtch.com/~rsc/regexp/regexp1.html)
* [Regular Expression Matching: the Virtual Machine Approach](https://swtch.com/~rsc/regexp/regexp2.html)
* [Regular Expression Matching in the Wild](https://swtch.com/~rsc/regexp/regexp3.html)
* [RE2 Syntax](https://github.com/google/re2/wiki/Syntax)
* [RE2 C++ source](https://github.com/google/re2)


